Sufficient Conditions for Reachability in Automata Networks with Priorities

### 2015-02-04 parts/sa.tex
remove comments
change order of subsections in section 3 (extraction; successive reachability; simultaneous reachability)
minor fixes
ls/nls -> ps/pps

\begin{definition}
\label{def:lcg}
Given a context $\ctx$ and a local state $\myp\in \Proc$,
the Local Causality Graph (LCG) for reachability under-approximation
$\cwB \DEF (\Bv, \Be)$
with
$\cwBNodes \subseteq \Proc \cup \Obj \cup \NSol \cup \sSol$
and
$\cwBEdges \subseteq \cwBNodes \times \cwBNodes$,
is the smallest graph such that:
\begin{enumerate}
\item
$\myp \in \cwBNodes$
\item
$a_i\in\cwBNodes\cap\Proc \Leftrightarrow \{ (a_i,\obj{a_j}{a_i}) \mid
a_j\in\ctx \vee a_j\in\cwBNodes\cap\Proc)
\}\subseteq\cwBEdges$
\label{lcg-ls-obj}
\item
$P\in\cwBNodes\cap\Obj \Leftrightarrow 
	\{ (P,\langle P,\abstr \zeta\rangle) \mid \zeta\in\BS(P) \}\subseteq\cwBEdges$
\label{lcg-obj-sol}
\item
$\langle P, nls \rangle\in\cwBNodes\cap\NSol \Leftrightarrow
	\{ (\langle P,nls\rangle, ls) \mid ls\in nls \}\subseteq\cwBEdges$
\label{lcg-sol-sync}
\item
$ls \in\cwBNodes\cap\sSol \Leftrightarrow
	\{ (ls, a_i) \mid a_i\in ls \}\subseteq\cwBEdges$
\label{lcg-sync-ls}
\item
$\obj{a_i}{a_j}\in \cwBNodes\cap\Obj \Rightarrow 
	\{(\obj{a_i}{a_j},\obj{a_k}{a_j}) \mid a_k\neq a_j,$
\label{lcg-conn}
\\
\hspace*{4cm}
$a_k\in \conn_{(\cwBNodes,\cwBEdges)}(\langle\obj{a_i}{a_j},\abstr\zeta\rangle),$
\\
\hspace*{4cm}
$\zeta\in\BS(\obj{a_i}{a_j}) \}\subset\cwBEdges$
\end{enumerate}
with $\abstr\zeta \DEF \{\PHhitter(\zeta_n) \mid n\in\indexes\zeta \}$.
\end{definition}

@@

The intuition is the following.
Given a state $s\subset\ctx$, we recursively refine the initial objective
(which is: reaching $\myp$ from the initial state) according to its children.
As the LCG is acyclic, such a recursion always terminates, and as all the objective nodes have at
least one child, it never gets stuck.
The refinement of an objective node $\obj{a_i}{a_j}$ acts as follows:
if the node has another objective node $\obj{a_k}{a_j}$ as child, we first refine the objective
$\obj{a_i}{a_k}$ (which, by construction, is necessarily in the LCG)
and then we refine the objective $\obj{a_k}{a_j}$.
If the objective node has only successors in $\Sol$ (bounce sequences), one is picked arbitrarily.
If $\langle P, nls\rangle$ is the chosen node,
by construction there exists $\zeta\in\BS(P)$ such that $\abstr\zeta = nls \in \powerset(\sSol)$.
If $\zeta = \emptyseq$ (for instance in the case when $a_i = a_j$), the recursion stops and we
continue to the next stage.
Otherwise, for each $n\in\indexes\zeta$,
for each $b_i \in\PHhitter(\zeta_n)$,
we refine the objective $\obj{b_j}{b_i}$ where $b_j$ is the state of $b$ in the current state.
By induction, $\obj{b_j}{b_i}$ is a child of $b_i$ in the LCG of \pref{def:lcg}.
Then, we know (by induction) that the current state of $b$ is $b_i$.
After having repeated this procedure for each $b_i\in\PHhitter(\zeta_n)$,
because all the nodes in $\sSol$, and in particular $\PHhitter(\zeta_n)$, are independent,
we know that all the local states in $\PHhitter(\zeta_n)$ are in the current state.
In addition, we know that the state of automaton $a$ has remained unchanged, otherwise
$\obj{a_i}{a_j}$ would have an objective child.
Hence, the action $\zeta_n$ is playable in the current state.
We can thus apply this action, modifying the state of $a$ to $a_j$ and continue to the next stage.
In the end, this recursive procedures builds a scenario from $s\subset\ctx$ to a state containing $\myp$.

@@

  \item In a $\sSol$ node with the current marking $m$:
    \begin{itemize}
      \item If $m \neq \emptyset$,
        descend in a child node $a_k \in \Proc$ arbitrarily chosen,
        so that $a_k \in m$ and carry on descending.
      \item If $m = \emptyset$,
        ascend in the parent $\sSol$ node.
        Let $\zeta$ be the current marking of this node.
        If $\zeta_1$ is playable in $s$:
          append $\zeta_1$ to the output,
          change the current state to $s \play \zeta_1$
          and mark this node with $\zeta_{2..|\zeta|}$.
        If $\zeta_1$ is not playable
        (which means that $\target{\zeta_1}$ has changed):
          go to the node $\PHobj{\PHget{s}{\PHsort(\bounce{\zeta_1})}}{\bounce{\zeta_1}} \in \Obj$
          and start descending.
    \end{itemize}
\end{itemize}

% * On applique les règles suivants de parcours où, à tout moment, s représente l'état courant, et l'état du parcours peut être « en descente » ou « en montée » :
% 
% - Dans un nœud a_k \in LS :
%   - Si en cours de descente, descendre dans le nœud s[a] />* a_k \in Obj.
%   - Si en cours de remontée, monter dans le nœud Sol parent et enlever l'élément a_k de son marquage.
% 
% - Dans un nœud a_j />* a_k \in Obj :
%   - Si en cours de descente, descendre dans un nœud pps \in NSol au choix. Marquer ce nœud avec une séquence \zeta \in BS tq \zeta^ = pps.
%   - Si en cours de remontée, monter dans le nœud LS ou Obj parent.
% 
% - Dans un nœud pps \in NSol (marqué par \zeta) :
%   - Si \zeta ≠ \emptyseq, descendre dans le nœud ps \in Sol tq ps = hitters(\zeta_1). Marquer le nœud ps par l'ensemble ps.
%   - Si \zeta = \emptyseq, alors remonter dans le nœud Obj parent et continuer la remontée.
% 
% - Dans un nœud Sol marqué par m :
%   - Si m ≠ \emptyset, alors descendre dans un des nœuds fils a_k \in LS tq a_k \in m.
%   - Si m = \emptyset, alors remonter dans le nœud NSol parent, marqué par \zeta. Si \zeta_1 est jouable, jouer cette action et marquer le nœud par \zeta_{2..|\zeta|}. Si \zeta_1 n'est pas jouable (c.-à-d. si target(\zeta_1) a changé), aller au nœud s[sort(bounce(\zeta_1))] />* bounce(\zeta_1) \in Obj et commencer une descente.

%%%%%%%%% Tnetative 1 %%%%%%%%%%%%%%%%%%%%%
% 
% \newcommand{\res}{\mathbf{res}}
% \newcommand{\osres}{\mathbf{OSres}}
% \newcommand{\autoBS}{\mathbf{autoBS}}
% \newcommand{\rec}{\mathbf{rec}}
% % \newcommand{\thisobj}[1]{P_{#1}}
% % \newcommand{\thisseq}[1]{\Theta_{#1}}
% \newcommand{\thisobj}{P}
% \newcommand{\thisseq}{\Theta}
% \newcommand{\R}{R}
% 
% First, $\osres$ is used to split the objective sequence $\w$ to solve,
% by considering successive objectives, in the fashion of \pref{thm:ordered-ua}.
% This function calls $\res$ which allows to solve an objective $P$
% by either re-target it (if $\gCont(\PHsort(P), P) \neq \emptyset$)
% of by searching for a solution to this objective in $\BS(P)$.
% This search may require to try several objective sequences $\zeta \in \BS(P)$.
% For each trial, the chosen objective sequence $\zeta$ is solved with $\rec$,
% which recursively created a scenario to play the whole sequence.
% This scenario is itself recursively created by combining
% smaller scenarios allowing to reach each required hitter of each action
% with the function $\thisseq$.
% Finally, $\thisseq$ calls $\res$ anew to reach the hitters
% that are necessarily contained in other automata.
% In the end, a scenario can be obtained by analysing all values of
% $\osres_s(\w)$ (\pref{def:concret})
% for all possible initial state $s \subseteq \ctx$ in the considered context,
% and all possible bounce sequences $\zeta \in \BS(P)$ that are chosen during the resolution.
% 
% \begin{definition}[$\osres : \PHl \times \OS \to \Sce$]
% \label{def:concret}
% For all $s \in \PHl$ and $\w \in \OS$, we define:
% \[
%   \osres_s(\w) \DEF
%     \begin{cases}
%       \emptyseq
%         & \text{ if $\w = \emptyseq$} \\
%       \res_s(\w_1) \concat
%       \osres_{s \play \res_s(\w_1)}(\w_2 \concat \dots \concat \w_{\card{\w}})
%         & \text{ otherwise} \\
%     \end{cases}
% \]
% and, for all $s \in \PHl$ and $P \in \Obj$:
% \[\res_s(P) \DEF
%   \begin{cases}
%     \varepsilon
%       & \text{if } \target{P} = \bounce{P} \\
%     \osres_s(P^Q \concat Q)
%       & \text{if } \exists Q \in \Obj, (P, Q) \in \Bee{\Obj}{\Obj} \\
%     \rec_s(\zeta)
%       & \text{otherwise, with } \zeta \in \BS(P)
%   \end{cases}
% \]
% with $P^Q = \PHobj{\PHtarget(P)}{\PHtarget(Q)}$.
% Furthermore:
% \begin{align*}
%   \forall \zeta \in \BS,
%   \rec_s(\zeta) &\DEF
%     \begin{cases}
%       \emptyseq
%         & \text{ if $\zeta = \emptyseq$} \\
%       \thisseq_s(\zeta_1) \concat \zeta_1 \concat
%       \rec_{s \play \thisseq_s(\zeta_1) \concat \zeta_1}
%       (\zeta_2 \concat \dots \concat \zeta_{\card{\zeta}})
%         & \text{ otherwise}
%     \end{cases} \\
%   \forall h \in \PHh,
%   \thisseq_s(h) &\DEF
%     \begin{cases}
%       \emptyseq
%         & \text{ if $\hitter{h} = \emptyset$} \\
%       \R_s(h_1) \concat \thisseq_{s \play \R_s(h_1)}(h_2 \concat \dots \concat h_{\card{h}})
%         & \text{ otherwise}
%     \end{cases} \\
%   \forall a_i \in \Proc,
%   \R_s(a_i) &\DEF \res_s(\PHobj{\PHget{s}{\PHsort(a_i)}}{a_i})
% \end{align*}
% and for all action $h \in \PHh$, $(h_i)_{i \in \segm{1}{\card{\hitter{h}}}}$
% is a sequence containing all hitters of $h$ in an arbitrary order.
% \end{definition}

%%%%%%%% Tentative 2 %%%%%%%%%%%%%%%%%%%%%%%
% % \newcommand{\resP}{\mathbf{\Obj res}}
% % \newcommand{\resOS}{\mathbf{\OS res}}
% \newcommand{\resP}{\mathsf{res}^\Obj}
% \newcommand{\resOS}{\mathsf{res}^\OS}
% 
% \newcommand{\res}{\mathbf{res}}
% \newcommand{\autoBS}{\mathbf{autoBS}}
% \newcommand{\rec}{\mathbf{rec}}
% 
% Pour tout objectif $P$ dans $\cwB$,
% on note $X_P$ une séquence d'objectifs obtenue en effectuant
% un parcours en profondeur d'abord de $\cwB$.
% Un tel parcours existe car $\cwB$ ne comporte pas de cycle.
% 
% $$\res_s(\w) = \resOS_s((X_{\w_i})_{i \in \indexes{\w}})$$
% 
% \[
%   \resOS_s(\w) \DEF
%     \begin{cases}
%       \emptyseq
%         & \text{ if $\w = \emptyseq$} \\
%       \resP_s(\w_1) \concat
%       \resOS_{s \play \resP_s(\w_1)}(\w_{2..\card{\w}})
%         & \text{ otherwise} \\
%     \end{cases}
% \]
% 
% $
% \resP_s(\PHobj{a_j}{a_k}) \DEF
%   \todo{resP}$
% 
% % $\resP_s(P) \DEF
% %     \begin{cases}
% %       \emptyseq
% %         & \text{ if $P$ is trivial} \\
% %       \res_s(\w_1) \concat
% %       \osres_{s \play \res_s(\w_1)}(\w_2 \concat \dots \concat \w_{\card{\w}})
% %         & \text{ otherwise} \\
% %     \end{cases}$

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

\begin{definition}
\label{def:lcg}
Given a context $\ctx$ and a local state $\myp\in \Proc$,
the Local Causality Graph (LCG) for reachability under-approximation
$\cwB \DEF (\Bv, \Be)$
with
$\cwBNodes \subseteq \Proc \cup \Obj \cup \NSol \cup \sSol$
and
$\cwBEdges \subseteq \cwBNodes \times \cwBNodes$,
is the smallest graph such that:
\begin{enumerate}
\item
$\myp \in \cwBNodes$
\item
$a_i\in\cwBNodes\cap\Proc \Leftrightarrow \{ (a_i,\obj{a_j}{a_i}) \mid
a_j \neq \myp \wedge (a_j\in\ctx \vee a_j\in\cwBNodes\cap\Proc)
\}\subseteq\cwBEdges$
\label{lcg-ls-obj}
\item
$P\in\cwBNodes\cap\Obj \Leftrightarrow 
	\{ (P,\langle P,\abstr \zeta\rangle) \mid \zeta\in\BS(P) \}\subseteq\cwBEdges$
\label{lcg-obj-sol}
\item
$\langle P, pps \rangle\in\cwBNodes\cap\NSol \Leftrightarrow
	\{ (\langle P,pps\rangle, ps) \mid ps\in pps \}\subseteq\cwBEdges$
\label{lcg-sol-sync}
\item
$ps \in\cwBNodes\cap\sSol \Leftrightarrow
	\{ (ps, a_i) \mid a_i\in ps \}\subseteq\cwBEdges$
\label{lcg-sync-ls}
\item
$\obj{a_i}{a_j}\in \cwBNodes\cap\Obj \Rightarrow 
	\{(\obj{a_i}{a_j},\obj{a_k}{a_j}) \mid a_k\neq a_j,$
\label{lcg-conn}
\\
\hspace*{4cm}
$a_k\in \conn_{(\cwBNodes,\cwBEdges)}(\langle\obj{a_i}{a_j},\abstr\zeta\rangle),$
\\
\hspace*{4cm}
$\zeta\in\BS(\obj{a_i}{a_j}) \}\subseteq\cwBEdges$
\end{enumerate}
with $\abstr\zeta \DEF \{\PHhitter(\zeta_n) \mid n\in\indexes\zeta \}$.
\end{definition}

@@

The intuition is the following.
Given a state $s\subset\ctx$, we recursively refine the initial objective
(which is: reaching $\myp$ from the initial state) according to its children.
As the LCG is acyclic, such a recursion always terminates, and as all the objective nodes have at
least one child, it never gets stuck.
The refinement of an objective node $\obj{a_i}{a_j}$ acts as follows:
if the node has another objective node $\obj{a_k}{a_j}$ as child, we first refine the objective
$\obj{a_i}{a_k}$ (which, by construction, is necessarily in the LCG)
and then we refine the objective $\obj{a_k}{a_j}$.
If the objective node has only successors in $\Sol$ (bounce sequences), one is picked arbitrarily.
If $\langle P, pps\rangle$ is the chosen node,
by construction there exists $\zeta\in\BS(P)$ such that $\abstr\zeta = pps \in \powerset(\sSol)$.
If $\zeta = \emptyseq$ (for instance in the case when $a_i = a_j$), the recursion stops and we
continue to the next stage.
Otherwise, for each $n\in\indexes\zeta$,
for each $b_i \in\PHhitter(\zeta_n)$,
we refine the objective $\obj{b_j}{b_i}$ where $b_j$ is the state of $b$ in the current state.
By induction, $\obj{b_j}{b_i}$ is a child of $b_i$ in the LCG of \pref{def:lcg}.
Then, we know (by induction) that the current state of $b$ is $b_i$.
After having repeated this procedure for each $b_i\in\PHhitter(\zeta_n)$,
because all the nodes in $\sSol$, and in particular $\PHhitter(\zeta_n)$, are independent,
we know that all the local states in $\PHhitter(\zeta_n)$ are in the current state.
In addition, we know that the state of automaton $a$ has remained unchanged, otherwise
$\obj{a_i}{a_j}$ would have an objective child.
Hence, the action $\zeta_n$ is playable in the current state.
We can thus apply this action, modifying the state of $a$ to $a_j$ and continue to the next stage.
In the end, this recursive procedures builds a scenario from $s\subset\ctx$ to a state containing $\myp$.

@@

  \item In a $\sSol$ node with the current marking $m$:
    \begin{itemize}
      \item If $m \neq \emptyset$,
        descend in a child node $a_k \in \Proc$ arbitrarily chosen,
        so that $a_k \in m$ and carry on descending.
      \item If $m = \emptyset$,
        ascend in the parent $\Sol$ node.
        Let $\zeta$ be the current marking of this node.
        If $\zeta_1$ is playable in $s$:
          append $\zeta_1$ to the output,
          change the current state to $s \play \zeta_1$
          and mark this node with $\zeta_{2..|\zeta|}$.
        If $\zeta_1$ is not playable
        (which means that $\target{\zeta_1}$ has changed):
          go to the node $\PHobj{\PHget{s}{\PHsort(\bounce{\zeta_1})}}{\bounce{\zeta_1}} \in \Obj$
          and start descending.
    \end{itemize}
\end{itemize}
